Stack Safety as a Security Property

### 2020-07-09 main.tex
done with pass on 3

\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states is the unit type, and its policy step
function is the constant unit function, thereby allowing all calls to the machine
step function. Its initialization function is defined on all states. This
corresponds to the lifting of our machine model to the policy setting. Machines
that fulfill a property by construction will also fulfill that property with the
null policy.

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler to
(e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
canaries \citep{Cowan+98}. These approaches are represented as policies that
accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
being produced by the compiler.) The set of policy states is again the unit type
and the step function constant. Property enforcement comes from the fact that
only ill-formed states can induce a trace that violates the property.

@@

\paragraph*{Capability machines}
%
Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies involving, for
example, compartmentalization or the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
mechanisms to policy states and step function would be a delicate operation, as
those are tightly coupled to the hardware design. However, separating the
enforcement aspects of a concrete policy (such as StkTokens, where policy state
would encompass details like the abstract call stack, the capability sealing
scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
the machine would be relatively straightforward.  \bcp{Not sure that any of
  this is comprehensible... :-(}

\paragraph*{Micropolicies}
%
Micropolicies are another major family of policies \citep{pump_oakland2015},
which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
Here the machine state is extended with tags paired with all components of the
system, and a monitor checks for each operation that the tags on the operands
obey a set of rules. One could treat tags directly as part of the machine
state, and build the rules into the step function. But local state
encapsulation properties as we frame them are not intuitive in their
interaction with tags or other additional enforcement state, so it is helpful
to separate tags for clarity. This also distinguishes a hardware fail-stop from
a software fail state.
\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
  policy? }

@@

To enforce this discipline, we define blessed header and return instruction
sequences. These are specific sequences of tagged instructions that are intended
to appear at the beginning and end of each function. This is the header
sequence:
%
\setcounter{pcctr}{1}
\[
  \begin{array}{l|l|l|l}
     & \mathit{Instruction} & \mathit{Tags} & \\
    \hline
    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
        {Store return address in stack}
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
        {Increment stack pointer by frame size}
  \end{array}
\]
%
And the return sequence:
%
\setcounter{pcctr}{1}
\[
  \begin{array}{l|l|l|l}
     & \mathit{Instruction} & \mathit{Tags} & \\
    \hline
    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
        {Load return address}
    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
        {Decrement stack pointer by frame size}
    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
        {Return to caller ($\jalr$)}
  \end{array}
\]

>>>>

This enforcement model is flexible enough to capture any kind of mechanism used
for dynamic enforcement of properties of interest, including the following
representative families of policies.
\apt{I don't really buy this, because we really depend on the ability to define
  variation and rollback. I think we should back off and say: this
  works well for micropolicies, and we conjecture that we can make it work
  for code-altering appraoches (pure software or capability).}

\paragraph*{Micropolicies}
%
Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
reference monitoring mechanism, which have been applied to stack safety by 
\citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step function
checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including 
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
(ordinary data) with base and bounds information (policy data),
and their use requires rewriting source code (e.g. to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate. \apt{But we conjecture it can be done...?}

%% Mapping the underlying protection
%% mechanisms to policy states and step function would be a delicate operation, as
%% those are tightly coupled to the hardware design. However, separating the
%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
%% would encompass details like the abstract call stack, the capability sealing
%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
%% the machine would be relatively straightforward.  \bcp{Not sure that any of
%%   this is comprehensible... :-(}

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler to
(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
canaries~\citep{Cowan+98}. \apt{not sure what to say here}
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

@@

%% \Paragraph*{Null policy}
%% %
%% The simplest possible policy is the neutral policy that carries no information
%% and does nothing. Its set of policy states is the unit type, and its policy step
%% function is the constant unit function, thereby allowing all calls to the machine
%% step function. Its initialization function is defined on all states. This
%% corresponds to the lifting of our machine model to the policy setting. Machines
%% that fulfill a property by construction will also fulfill that property with the
%% null policy.

@@

To enforce this discipline, we define blessed header and return instruction
sequences. These are specific sequences of tagged instructions that are intended
to appear at the beginning and end of each function:

\[
  \begin{array}{c|c|c}
    {\sw ~ \rsp ~ \rra ~ 1} & {\tagHa, \tagInstr} & \text{Store RA in stack} \\
    {\addi ~ \rsp ~ \rsp ~ 2} & {\tagHb, \tagInstr} & \text{Increment SP by frame size}
  \end{array}
  \]

\[
  \begin{array}{c|c|c}
    {\lw ~ \rra ~ \rsp ~ -1} & {\tagRa, \tagInstr} & \text{Load RA} \\
    {\addi ~ \rsp ~ \rsp ~ -2} & {\tagRb, \tagInstr} & \text{Decrement SP by frame size} \\
    {\jalr ~ \rra ~ \rra} & {\tagRc, \tagInstr} & \text{Jump back}
  \end{array}
\]
